---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r echo=TRUE}
.libPaths("./library") # set library path to my own. You may change this to your own library
library("dplyr") # load dplyr package
ac_data <- read.csv("activity.csv")
head(ac_data)
actdata <- tbl_df(ac_data)
```


## What is total number of steps taken per day?
The total number of steps taken per day is:
```{r echo=TRUE}
total_steps <- sapply(split(actdata$steps,actdata$date),sum,na.rm=T)
```

### Histgram of the total number of steps taken per day.
```{r echo=TRUE, fig.width=6,fig.height=4}
hist(total_steps,xlab="Day",ylab="Total number of stpes per day")
```

### The mean and median of the total number of stpes per day are:
```{r echo=TRUE}
mean(total_steps)
median(total_steps)
```

## What is the average daily activity pattern?
```{r echo=TRUE}
fivemin_act <- group_by(actdata,interval)
fivemin_act <- summarise(fivemin_act,mean=mean(steps,na.rm=T))
plot(fivemin_act,type="l")
```

## Imputing missing values
```{r echo=TRUE}
sum(is.na(actdata$steps)) ## Calculate the total number of missing value
```
There are `r sum(is.na(actdata$steps))` missing values. For simplicity, they are all filled by 15.

``` {r echo=TRUE}
newdata <- mutate(actdata,steps = replace(steps, is.na(steps), 15))
```
The histgram of the total number of steps taken each day after filled NAs.
```{r echo=TRUE, fig.width=6,fig.height=4}
new_total_steps <- group_by(newdata,date) %>% summarise(total=sum(steps))    
hist(new_total_steps$total,xlab="Day",ylab="Total number of stpes per day")
```


The mean of the new dataset is:

```r {echo=TRUE}
mean(new_total_steps$total)
```

The median is:

```r {echo=TRUE}
median(new_total_steps$total)
```

## Are there differences in activity patterns between weekdays and weekends?
